### 2015-06-12 text/ms.Rmd
Edit introduction and methods; add to results

We often have multiple models of ecological systems. For example, models might
incorporate different data types, assume alternate population dynamics, or make
contrasting assumptions about the starting state of a system. These models may
suggest different conclusions about population status and often times we do not
know which model is best. How can we reconcile multiple models to make robust
management decisions about ecological resources?

Ensemble models, which combine the output from multiple models, are one
potential solution to this problem. A simple ensemble model might take the
average of multiple model outputs, perhaps weighted by some goodness of fit. A
complex ensemble model might also take into account the interactions between
model outputs and additional covariates. Ensemble models are popular in other
fields including public health [e.g. @thomson2006], agriculture [e.g.
@cantelaube2005], climate science [e.g. @murphy2004; @tebaldi2007;
@pierce2009], and machine learning [e.g. @dietterich2000]. We think they have
tremendous opportunity in ecological resource management.

In fisheries science, a common problem is estimating the status of an exploited
fish population. For the majority of fish stocks, we have limited biological
information and stock status isn't known [@fao2014]. In recent years, a number
of methods have been proposed to derive population status based on the limited
information that is frequently available: fisheries catch time series and basic
knowledge about the productivity of the species. Recently, @rosenberg2014
investigated the ability for these models to estimate population status through
a large-scale simulation model. A key finding was that these models frequently
disagree on population status and any one model has relatively poor predictive
ability on average (e.g. Fig. \ref{fig:motivate}).

@@

Here, we develop ensemble models to estimate the population status of exploited
fish populations. We explore a variety of ensemble approaches applied to both
simulated and real-world fish stocks and compare predictive performance. In
developing these models we highlight the types of problems one encounters when
forming and evaluating ensemble models. We find that ensembles generally
provide improved predictive performance than any single model---especially when
considered across multiple dimensions of predictive ability. Furthermore, the
output from ensemble models can provide better mechanistic understanding of how
individual models perform.

@@

We tested the ability of ensemble models to improve estimates of population
status when applied to both a large simulated dataset and a global database of
assessed stock status. Here we describe the datasets, individual models of
population status, and ensemble models to combine those estimates. We then
describe how we evaluated the ability of the various models to estimate
population status.

@@

- *mPRM* (modified panel regression model) is a modified version of the
panel-regression model used in @costello2012. Unlike the other models
considered, mPRM is empirical and not mechanistic --- it fits a linear model to
a series of characteristics of the catch time series and stock with \bbmsy\\ as
the response. The model used in this paper and @rosenberg2014 is modified from
the original in that it simplifies the life-history categories to be applicable
to simulated data.
<!--TODO: what are we doing with the RAM stocks again?-->

<!--TODO: build in fig:didactic -->

@@

Ensemble methods allow us to incorporate additional covariates into our models
and potentially leverage interactions of how these covariates combine with the
main individual model estimates to estimate population status. Examples might
be life-history characteristics, additional information on exploitation
patterns, or additional statistical properties of the data going into the
individual models. For simplicity, and to allow us to apply models developed
with the simulated dataset to the real-world dataset, we added only one set of
additional covariates: spectral properties of the catch time series itself.
Spectral analysis decomposes a time series into the frequency domain and
provides a useful means of describing the statistical properties of the catch
series that is independent of time series length and magnitude of catch. We fit
spectral models to the scaled catch time series (divided by maximum catch) with
the `spec.ar` function in \textsf{R} and recorded representative short- and
long-term spectral densities of 0.05 and 0.2, which correspond to 20- and
5-year cycles.

@@

The individual models we seek to combine with ensemble models provide time
series of \bbmsy\\ ($\hat{b}_\mathrm{CMSY}$, $\hat{b}_\mathrm{COM-SIR}$, etc.).
Therefore, we can use ensemble models to estimate multiple properties of these
status time series. Here, we focus on two values: the mean and slope of
\bbmsy\\ in the last 5 years. Below, we refer to these quantities collectively
as $\hat{b}_\mathrm{ensemble}$. Together, these quantities address the current
state and trajectory of status, both of which may be of management and
conservation interest. To avoid undue influence of the end points of the time
series on the calculated slope, we measured the slope as the Theil-Sen
estimator of median slope [@theil1950].

@@

$$
\hat{\theta} = 
\mathrm{argmin}_\theta \left( L(b \, | \, \hat{b}_{\mathrm{ensemble}}) \right)
$$
$$
\hat{b}_\mathrm{ensemble} = 
f \left( \hat{b}_\mathrm{CMSY}, \, \hat{b}_\mathrm{COM-SIR}, \,
\hat{b}_\mathrm{SSCOM}, \, \hat{b}_\mathrm{mPRM}, \,
S(0.2), \, S(0.5), \, | \, \theta \right)
$$

\noindent where $L$ is a generalized loss function, $f$ is some generalized
regression function, $\hat{b}_\mathrm{CMSY}$ is the prediction from the CMSY
model etc. (Table \ref{tab:predictors}), $S(n)$ represents the spectral density
at frequency $n$ of the scaled catch series, $b$ is the true value the model is
fitted to, and $\theta$ is the set of parameters for the regression function.
For all ensemble models of mean \bbmsy\\, which is bounded at zero, we fit the
models in log space and exponentiated the predictions to derive a median
estimate of \bbmsy\\. For the estimates of \bbmsy\\ slope we fit ensemble
models on the natural untransformed scale.

We calculated the average ensemble models as the geometric mean of the
individual model median \bbmsy\\ estimates and the mean of the individual model
\bbmsy\\ slopes. We fit the linear model ensemble with second-order
interactions. We fit two machine learning ensemble models based on regression
trees. Regression trees sequentially determine what value of a predictor best
splits the response data into two 'branches' based on a specified loss
function. In random forests, a series of regression trees are built on a random
subset of the data and random subset of the covariates of the model (REF). In
generalized boosted models (GBMs), each subsequent model is fit to the
residuals from the previous model; data points that are fit poorly in a given
model are given more weight in the next model [@elith2008]. Random forests and
GBMs can provide strong predictive performance and fit highly non-linear
relationships (REFs). We fit random forest models with the **randomForest**
package [@liaw2002] for \textsf{R} with the default argument values. We fit
boosted regression tree models with the **GBM** package [@ridgeway2015] for
\textsf{R}. Based on cross-validation with the **caret** package (Supporting
Materials, Fig. SX), we fit GBMs with 2000 trees, an interaction depth of $6$,
a learning rate (shrinkage parameter) of $0.01$, and all other arguments at
their default values.

@@

A critical component to any predictive modelling exercise is to evaluate the
performance of a model new data [@hastie2009]. We used repeated three-fold
cross validation to test predictive performance: we randomly divided the
dataset into three sets, built ensemble models on two-thirds of the data, and
evaluated predictive performance on the remaining third of the data. We
repeated this across each of the three splits and then repeated the whole
procedure XX times to account for bias that may result from any one set of
validation splits. Since the dynamics of simulated populations differing only
in the magnitude of observation error around catch, magnitude of recruitment
variability, or stochastic draw of these values are likely similar, we grouped
these characteristics together in the cross-validation process---we never
trained and tested on models differing only in these characteristics.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of bias,
precision, accuracy (a combination of bias and precision), or the ability to
correctly rank or correlate across populations. Here, we measure proportional
error (PE; also called 'relative error') defined as $(\hat{b} - b)/b$. We
summarize median proportional error to measure bias, median absolute
proportional error to measure accuracy, and Spearman's rank-order correlation
to measure the ability to correctly rank populations.

@@

machine learning models and linear models outperformed all individual models in terms of accuracy (MAPE) and Spearman's rank-order correlation across stocks

@@

For example, the linear model ensemble suggests that if both mPRM and SSCOM predict high \bbmsy\\, the true \bbmsy\\ is also high (Fig. \ref{fig:lm-coefs})

The GBM model, also indicates a similar relationship, but suggests a non-linear relationship between these individual models in predicting \bbmsy\\ (Fig. \ref{fig:partial-2d-sim}XX)

@@

Ensemble models can exploit the best performance characteristics of individual models. 
For example, if SSCOM predicts a low \bbmsy\\ (less than about 0.X), true \bbmsy\\ is probably low (Fig. \ref{fig:hexagon}c).
But, SSCOM provides less predictive ability on its own when it predicts higher \bbmsy\\. The machine learning ensemble models can exploit this (Fig. \ref{fig:partial-sim}XX). On its own, CMSY contributes the most to predicting \bbmsy\\ when it makes predictions of \bbmsy\\ over one.

@@

Some individual models performed nearly as well as the enembles in one or two performance dimensions: e.g. SSCOM does nearly as well as the ensembles at predicting the slope of \bbmsy\\ (Fig. \ref{fig:performance-sim-slope}). But no individual method is consistently nearly as good as the ensembles. SSCOM does more poorly than the ensembles and other individual models in terms of accuracy, rank-order correlation, and bias at estimating the mean \bbmsy\\ for both the simulated and stock-assessment datasets (Fig. \ref{fig:performance).

>>>>

There are often multiple ecological models we can apply to a system to assess
population status and trajectory. Models might incorporate different data
types, assume alternate population dynamics, or make contrasting assumptions
about the starting state of a system. These models may suggest different
conclusions about population status and trajectory and we often do not know
which model is best. How can we reconcile multiple models to make robust
management decisions about ecological resources?

Ensemble methods, which combine the output from multiple models as predictors
in a new 'ensemble' model, are one solution to this problem. A simple ensemble
model might take the average of multiple model outputs, perhaps weighted by
some goodness of fit. A complex ensemble model might also take into account the
interactions between model outputs and additional covariates. Ensemble models
are popular in other fields including public health [e.g. @thomson2006],
agriculture [e.g. @cantelaube2005], climate science [e.g. @murphy2004;
@tebaldi2007; @pierce2009], and machine learning [e.g. @dietterich2000]. We
think they have tremendous opportunity in ecological resource management.

In fisheries science, a common problem is estimating the status and trajectory
of an exploited fish population. For the majority of fish stocks, we have
limited biological information and stock status isn't known [@fao2014]. In
recent years, a number of methods have been proposed to derive population
status based on the limited information that is broadly available: fisheries
catch time series and basic knowledge about the productivity of the species.
Recently, @rosenberg2014 investigated the ability for these models to estimate
population status through a large-scale simulation model. A key finding was
that these models frequently disagree on population status and any one model
has relatively poor predictive ability on average (e.g. Fig.
\ref{fig:motivate}).

@@

Here, we develop ensemble models to estimate the population status and
trajectory of exploited fish populations. We explore a variety of ensemble
approaches applied to both simulated and real-world fish stocks and compare
predictive performance. In developing these models we highlight the types of
challenges one encounters when forming and evaluating ensemble models. We find
that ensembles generally provide improved predictive performance than any
single model---especially when considered across multiple dimensions of
predictive ability and across multiple datasets. Furthermore, the output from
ensemble models can provide better mechanistic understanding of how individual
models of status and trajectory perform.

@@

We tested the ability of ensemble models to improve estimates of population
status and trajectory when applied to both a large simulated dataset and a
global database of assessed stock status. Here we describe the datasets,
individual models of population status, and ensemble models to combine those
estimates. We then describe how we evaluated the ability of the various models
to estimate population status.

@@

- *mPRM* (modified panel regression model) is a modified version of the
  panel-regression model used in @costello2012. Unlike the other models
  considered, mPRM is empirical and not mechanistic --- it uses the RAM Legacy
  Stock Assessment database to fit a linear model to a series of
  characteristics of the catch time series and stock with stock-assessed
  \bbmsy\\ as the response. The model used in this paper and @rosenberg2014 is
  modified from the original in that it condenses the life-history categories
  into three categories to match the simulated dataset. 
  <!--TODO: what are we doing with the RAM stocks again?-->
  <!--TODO: bias adjustment?-->
  <!--TODO: build in fig:didactic -->

@@

Ensemble methods allow us to incorporate additional covariates into our models
and potentially leverage interactions between these covariates and individual
model predictions. Additional covariates could be, for example, life-history
characteristics, information on exploitation patterns, or statistical
properties of the data. For simplicity, and to allow us to apply models
developed with the simulated dataset to the real-world dataset, we added only
one set of additional covariates: spectral properties of the catch time series.
Spectral analysis decomposes a time series into the frequency domain and
provides a means of describing the statistical properties of the catch series
that is independent of time series length and magnitude of catch. We fit
spectral models to the scaled catch time series (catch divided by maximum
catch) with the `spec.ar` function in \textsf{R} and recorded representative
short- and long-term spectral densities of 0.05 and 0.2, which correspond to
20- and 5-year cycles.

@@

The individual models we seek to combine with ensemble models provide time
series of \bbmsy\\. Therefore, we can use ensemble models to estimate multiple
properties of these status time series. Here, we focus on two values: the mean
and slope of \bbmsy\\ in the last 5 years. Together, these quantities address
the recent state and trajectory of status, both of which may be of management
and conservation interest. To avoid undue influence of the end points of the
time series on the calculated slope, we measured the slope as the Theil-Sen
estimator of median slope [@theil1950].

@@

$$
\hat{\theta} = 
\argmin_\theta \left( L(b \, | \, \hat{b}_{\mathrm{ensemble}}) \right)
$$
$$
\hat{b}_\mathrm{ensemble} = 
f \left( \hat{b}_\mathrm{CMSY}, \, \hat{b}_\mathrm{COM-SIR}, \,
\hat{b}_\mathrm{SSCOM}, \, \hat{b}_\mathrm{mPRM}, \,
S(0.2), \, S(0.5), \, | \, \theta \right)
$$

\noindent where $L$ is a generalized loss function, $f$ is some generalized
regression function, $\hat{b}_\mathrm{CMSY}$ is the prediction from the CMSY
model etc. (Table \ref{tab:predictors}), $S(0.2)$ and $S(0.05)$ represent the
spectral density of the scaled catch series at frequencies of 0.2 and 0.05, $b$
is the true value the model is fitted to, and $\theta$ is the set of parameters
for the regression function. For all ensemble models of mean \bbmsy\\, which is
bounded at zero, we fit the models in log space and exponentiated the
predictions to derive a median estimate of \bbmsy\\. For the estimates of
\bbmsy\\ slope, we fit ensemble models on the natural untransformed scale.

We calculated the average ensemble models as the geometric mean of the
individual model \bbmsy\\ estimates and the mean of the individual model
\bbmsy\\ slopes. We fit the linear model ensemble with second-order
interactions. We fit two machine learning ensemble models based on regression
trees. Regression trees sequentially determine what value of a predictor best
splits the response data into two 'branches' based on a specified loss
function. In random forests, a series of regression trees are built on a random
subset of the data and random subset of the covariates of the model (REF). In
generalized boosted models (GBMs), each subsequent model is fit to the
residuals from the previous model; data points that are fit poorly in a given
model are given more weight in the next model [@elith2008]. Random forests and
GBMs can provide strong predictive performance and fit highly non-linear
relationships [e.g. @elith2008; @hastie2009]. We fit random forest models with
the **randomForest** package [@liaw2002] for \textsf{R} with the default
argument values. We fit boosted regression tree models with the **GBM** package
[@ridgeway2015] for \textsf{R}. Based on cross-validation with the **caret**
\textsf{R} package, we fit GBMs with 2000 trees, an interaction depth of $6$, a
learning rate (shrinkage parameter) of $0.01$, and all other arguments at their
default values.

@@

A critical component to any predictive modelling exercise is to evaluate the
performance of a model on new data [@hastie2009]. We used repeated three-fold
cross validation to test predictive performance: we randomly divided the
dataset into three sets, built ensemble models on two-thirds of the data, and
evaluated predictive performance on the remaining third. We repeated this
across each of the three splits and then repeated the whole procedure 50 times
to account for bias that may result from any one set of validation splits.
Since the dynamics of simulated populations differing only in the stochastic
draw of random values are likely similar, we grouped these stocks (10 per
factorial cell) in the cross-validation process into either the training or
testing split.

Predictive performance can be evaluated with metrics that represent a variety
of modelling goals. For continuous response variables such as the mean and
slope of population status, performance metrics often measure some form of bias,
precision, accuracy (a combination of bias and precision), or the ability to
correctly rank or correlate across populations (REFs). Here, we measure proportional
error (PE; also called 'relative error') defined as $(\hat{b} - b)/b$. We
summarize median proportional error to measure bias, median absolute
proportional error to measure accuracy, and Spearman's rank-order correlation
to measure the ability to correctly rank populations.

@@

non-linear machine learning models were best able to predict higher \bbmsy\\ values (Fig. \ref{fig:hexagon}g, h) by exploiting interactions between the spectral density predictors and the individual model predictions (Fig. \ref{fig:partial-2d-sim})

machine learning and linear ensemble models outperformed all individual models
in terms of accuracy (MAPE) and Spearman's rank-order correlation across stocks
at predicting mean \bbmsy\\ with the simulated dataset of known status (Fig.
\ref{fig:performance}a)

the level of correlation across stocks and accuracy were both lower when
predicting full-stock assessement \bbmsy\\ than when predicting true status in
the simulated dataset (Fig. \ref{fig:performance}a vs. b)

@@

For example, the linear model ensemble suggests that if both mPRM and SSCOM
predict high \bbmsy\\, the true \bbmsy\\ is also high (Fig. \ref{fig:lm-coefs})
The GBM model, also indicates a similar relationship, but suggests a non-linear
relationship between these individual models in predicting \bbmsy\\ (Fig.
\ref{fig:partial-2d-sim}l)

@@

Ensemble models can exploit the best performance characteristics of individual
models. For example, if SSCOM predicts a low \bbmsy\\ (less than about 0.X),
true \bbmsy\\ is probably low (Fig. \ref{fig:hexagon}c). But, SSCOM provides
less predictive ability on its own when it predicts higher \bbmsy\\. The
machine learning ensemble models can exploit this (Fig.
\ref{fig:partial-sim}c). On its own, CMSY contributes the most to predicting
\bbmsy\\ when it makes predictions of \bbmsy\\ between about 1.0 and 1.4 (Fig.
\ref{fig:partial-sim}a)

@@

Some individual models performed nearly as well as the ensembles in one or two
performance dimensions: e.g. SSCOM does nearly as well as the ensembles at
predicting the slope of \bbmsy\\ (Fig. \ref{fig:performance-sim-slope}). But no
individual method is consistently nearly as good as the ensembles. SSCOM does
more poorly than the ensembles and other individual models in terms of
accuracy, rank-order correlation, and bias at estimating the mean \bbmsy\\ for
both the simulated and stock-assessment datasets (Fig. \ref{fig:performance).
